---
title: Bayesian Graphical Lasso for Estimating Brain Networks
author: Liam Hanson
output: 
  beamer_presentation:
    theme: "default"
    colortheme: "dolphin"
    fonttheme: "structurebold"
bibliography: references.bib
---

# Motivation

We are interested in viewing the brain as a graph i.e. a set of nodes, say regions of the brain, and edges, the connections between these edges

```{r, out.width="60%", fig.cap= "[@Hendricksen2015VisualizingDB]"}
#| echo: FALSE
knitr::include_graphics(
  here::here("figures", "13-Figure1.6-1.png")
)
```

Different brains have different network structures and therefore different local and global properties of these networks

# Motivation

In general, healthy brains have a so-called “small-world” or “scale-free” structure.

Each node has few neighbors but it is easy for one group of nodes or sub-network to talk to another (six degrees of separation). 

This implies the existence of hub nodes,
nodes through which other areas of the network communicate.


# Motivation 

```{r, out.width="60%", fig.cap= "[@Bullmore2009]"}
#| echo: FALSE
knitr::include_graphics(
  here::here("figures", "41583_2009_Article_BFnrn2575_Fig3_HTML.png")
)
```

A contrasting example is in people with Schizophrenia. These individuals
tend to score lower on these small-world metrics. 
 
# Graph Estimation

We don't observe these graphs directly

```{r, out.width="60%", fig.cap="[@Buckner2013]"}
#| echo: FALSE
knitr::include_graphics(
  here::here("figures", "41593_2013_Article_BFnn3423_Fig1_HTML.png")
)
```

Instead, we have time series data for each voxel or region of interest in the
brain where the measured quantity is the so-called BOLD signal.

We will use the associations between these time series to estimate our graph.

# Graph Estimation

A graph can be represented as an adjacency matrix: a symmetric matrix $A$ with
$a_{ij} = 1$ if an edge is present between node $i,j$ and $a_{ij} = 0$ otherwise.

For our purposes, we assume that at time step (i.e. scan) $t$, the value of the 
BOLD signal across all $D$ nodes $Y_D \sim MVN(0, \Sigma)$.

Fact: The precision matrix $\Omega = \Sigma^{-1}$ has the property that its entries
$\omega_{ij}$ encode the partial correlation between node $i$ and node $j$. 

If $\omega_{ij} = 0$, these nodes are conditionally independent.

# Graphical Lasso

- To capture this sparsity, something like LASSO (Tibshirani 1996)
could work.

- We could fit successive pairwise LASSO's and regress the residuals to obtain 
partial correlations, but this isn't ideal

- Instead, let's maximize the penalized multivariate gaussian log-likelihood
with respect to the precision matrix:

$$\textrm{log (det }\Omega) - \textrm{tr}(\frac{S}{n}\Omega) - \rho||\Omega||_1$$
- where $S$ is the sample covariance matrix and $||\Omega||_1$ is the entry-wise
$l_1$-norm of $\Omega$. 

# Graphical Lasso

- This is the approach of @friedman2007sparseinversecovarianceestimation

- At a high level, it recursively fits a modified LASSO in a blockwise fashion
using sub-gradient methods (gets around non-differentiability of $l_1$ norm.)

- The focus of my project was a Bayesian approach to the Graphical Lasso by 
@10.1214/12-BA729 to estimate the aforementioned graphs

# Bayesian Graphical Lasso 

We have the likelihood and prior
$$p(\textbf{y}_t | \Omega) = \textrm{N}(\textbf{y}_t|0, \Omega^{-1} )$$
$$p(\Omega|\lambda) = C^{-1} \prod_{i<j}\textrm{DE}(\omega_{ij}|\lambda)\prod_{i=1}^p
\textrm{Exp}(\omega_{ij}|\frac{\lambda}{2})\cdot1_{\Omega\in M^+}$$

where $M^+$ is the space of positive-definite matrices. 











